Multi-Domain Adaptation in Neural Machine Translation with Dynamic Sampling Strategies

### 2022-03-31 2022/eamt/eamt-latex/eamt22.tex
10 pages - sharp

We conventionally define a domain $d$ as a distribution $\mathcal{D}_d(x)$ over some feature space $\mathcal{X}$ that is shared across domains \cite{Pan10asurvey}: in machine translation, $\mathcal{X}$ is the representation space for input sentences; each domain corresponds to a specific source of data, and may differ from other data sources in terms of textual genre, thematic content \cite{Chen16guided,Zhang16topicinformed}, register \cite{Sennrich16politeness}, style \cite{Niu18multitask}, etc. Translation in domain $d$ is formalized by a translation function $h_d(y|x)$ pairing sentences in a source language with sentences in a target language $y \in \mathcal{Y}$. $h_d$ is usually assumed to be deterministic (hence $y = h_d(x)$) but may differ across domains.

@@

We experiment with translation from English into French in 6~domains, corresponding to the following data sources: the UFAL Medical corpus V1.0 (\domain{med})\footnote{\url{https://ufal.mff.cuni.cz/ufal_medical_corpus}. We only use the in-domain (medical) subcorpora: PATR, EMEA, CESTA, ECDC.}; the European Central Bank corpus (\domain{bank}) \cite{Tiedemann12parallel}; the JRC-Acquis Communautaire corpus (\domain{law}) \cite{Steinberger06acquis}; documentations for KDE, Ubuntu, GNOME and PHP from the Opus collection \cite{Tiedemann09news}, merged in a \domain{it}-domain; TedTalks (\domain{talk}) \cite{Cettolo12wit}, and the Koran (\domain{rel}). Additional experiments use the News Commentary corpus (\domain{news}). Most corpora are available from the Opus website.\footnote{\url{http://opus.nlpl.eu}} These corpora were deduplicated and tokenized with in-house tools; statistics are in Table~\ref{tab:Corpora}. To reduce the number of types, we use Byte-Pair Encoding \cite{Sennrich16BPE} with 30,000 merge operations on a corpus containing all sentences in both languages.\fyDone{Add \# number of tokens, also specificity ?}%
%
We randomly select in each corpus a development and a test set of 1,000 lines and keep the rest for training. Validation sets are used to chose the best model according to the average BLEU score \cite{Papineni02bleu}.\footnote{We use truecasing and \texttt{sacrebleu} \cite{Post18call}.}\fyDone{A word about meta-parameter settings} Significance testing is performed using bootstrap resampling \cite{Koehn04statistical}, implemented in compare-mt\footnote{\url{https://github.com/neulab/compare-mt}} \cite{Neubig19compare-mt}. We report significant differences at the level of $p=0.05$.\fyDone{Fix correct p value}

@@

Domain adaptation is an old problem that has been studied from many angles, both for SMT and NMT. A survey of supervised and unsupervised DA for NMT is in \cite{Chu18asurvey}, where the authors distinguish between data-centric and model-centric DA, a view also adopted in the recent survey of \newcite{Saunders21asurvey}. Our approach to DA in this paper falls under the former category. We refer readers interested in DA to these papers.

>>>>

We conventionally define a domain $d$ as a distribution $\mathcal{D}_d(x)$ over some feature space $\mathcal{X}$ that is shared across domains \cite{Pan10asurvey}: in machine translation, $\mathcal{X}$ is the representation space for input sentences; each domain corresponds to a specific source of data, and may differ from other data sources in terms of textual genre, thematic content \cite{Chen16guided,Zhang16topicinformed}, register, style \cite{Niu18multitask}, etc. Translation in domain $d$ is formalized by a translation function $h_d(y|x)$ pairing sentences in a source language with sentences in a target language $y \in \mathcal{Y}$. $h_d$ is usually assumed to be deterministic (hence $y = h_d(x)$) but may differ across domains.

@@

We experiment with translation from English into French in 6~domains, corresponding to the following data sources: the UFAL Medical corpus V1.0 (\domain{med})\footnote{\url{https://ufal.mff.cuni.cz/ufal_medical_corpus}. We only use the in-domain (medical) subcorpora: PATR, EMEA, CESTA, ECDC.}; the European Central Bank corpus (\domain{bank}); the JRC-Acquis Communautaire corpus (\domain{law}) \cite{Steinberger06acquis}; documentations for KDE, Ubuntu, GNOME and PHP from the Opus collection, merged in a \domain{it}-domain; TedTalks (\domain{talk}) \cite{Cettolo12wit}, and the Koran (\domain{rel}). Additional experiments use the News Commentary corpus (\domain{news}). Most corpora are available from the Opus website \cite{Tiedeman12parallel}.\footnote{\url{http://opus.nlpl.eu}} These corpora were deduplicated and tokenized with in-house tools; statistics are in Table~\ref{tab:Corpora}. To reduce the number of types, we use Byte-Pair Encoding \cite{Sennrich16BPE} with 30,000 merge operations on a corpus containing all sentences in both languages.\fyDone{Add \# number of tokens, also specificity ?}%
%
We randomly select in each corpus a development and a test set of 1,000 lines and keep the rest for training. Validation sets are used to chose the best model according to the average BLEU score \cite{Papineni02bleu}.\footnote{We use truecasing and \texttt{sacrebleu} \cite{Post18call}.}\fyDone{A word about meta-parameter settings} Significance testing is performed using bootstrap resampling \cite{Koehn04statistical}, implemented in compare-mt\footnote{\url{https://github.com/neulab/compare-mt}} \cite{Neubig19compare-mt}. We report significant differences at the level of $p=0.05$.\fyDone{Fix correct p value}

@@

Domain adaptation is an old problem that has been studied from many angles, both for SMT and NMT. A survey of supervised and unsupervised DA for NMT is in \cite{Chu2017comparison}, where the authors distinguish between data-centric and model-centric DA, a view also adopted in the recent survey of \newcite{Saunders21asurvey}. Our approach to DA in this paper falls under the former category. We refer readers interested in DA to these papers.
